---
title: "Signature Greffe"
author: "LIPINSKI Boris"
date: '2022-09-01'
output:
  rmdformats::robobook:
    code_folding: hide
    fig.width: 9
    fig.height: 7
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(cache.lazy = F)
path <- "/media/boris/bae7e14e-21e5-48b8-80d6-f94583367b83"
annotations <- read.csv(paste0(path,"/Rapport-RNAseq/annotation.csv"))
source(file = paste0(path,"/Rapport-RNAseq/Functions.R"))
options(digits = 3)

condition <- "Condition" ; item1 <- "Pregreffe" ; item2 <- "Excipient" ; FC <- 1.189207 #FC <- 2.7185
load(file = paste0(path,"/Flinovo/analyse_meta/All_Pre-greffe-Excipient_B.RData"))
singlet@meta.data$Condition <- stringr::str_replace(singlet@meta.data$Condition, "Pré-greffe", "Pregreffe")
DefaultAssay(singlet) <- "RNA" ; singlet <- noig(singlet) ; Idents(singlet) <- condition ; sc <- PB(singlet, condition)
```

```{r echo=F}
sc.EdgeR0 <- EdgeR.Workflow(counts = sc$counts, metadata = sc$metadata, group = sc$metadata$Condition, FDR = 0.05,
                           protocol = ~ Sample+Condition, FC = FC, custom = F, treat = T, glm = T, exact = F)
```

# DEG {.tabset}

Cellule B uniquement + FC = 0.25
 

## Pseudo-Bulk - les databases (union DB papiers) {.tabset}
```{r echo=F, message=FALSE}
load("/home/boris/Bureau/Flinovo/script/Signature/Greffe/DB/diss.gene.union")
singlet@assays$RNA@data <- singlet@assays$RNA@data[rownames(singlet@assays$RNA@data)[!(rownames(singlet@assays$RNA@data) %in% diss.gene.union)],];
singlet@assays$RNA@scale.data <- singlet@assays$RNA@scale.data[rownames(singlet@assays$RNA@scale.data)[!(rownames(singlet@assays$RNA@scale.data) %in% diss.gene.union)],];
singlet@assays$RNA@counts <- singlet@assays$RNA@counts[rownames(singlet@assays$RNA@counts)[!(rownames(singlet@assays$RNA@counts) %in% diss.gene.union)],];

sc <- PB(singlet, condition)
sc.EdgeR <- EdgeR.Workflow(counts = sc$counts, metadata = sc$metadata, group = sc$metadata$Condition, FDR = 0.05,
                           protocol = ~ Sample+Condition, FC = FC, custom = F, treat = T, glm = T, exact = F)
```


### Liste {.tabset}

#### treat
```{r echo=F}
DE.Greffe.PB2 <- result(sc.EdgeR$treat$result) ; DT.table(DE.Greffe.PB2)
DE.Greffe.PB2.raw <- sc.EdgeR$treat$result
save(DE.Greffe.PB2.raw, file = "../Greffe/DE.Greffe.PB2.raw.Rdata")
```

#### Glm
```{r echo=F}
DE.Greffe.PB.Glm2 <- result(sc.EdgeR$glm$result) ; DT.table(DE.Greffe.PB.Glm2)
```

### Figures {.tabset}

#### Clustering
```{r echo=F, warning=F, fig.width=9, fig.height=8}
Clustering(sc.EdgeR$treat)
```

#### PCA
```{r echo=F}
PCA_edger(sc.EdgeR$treat, position = 'right')
```

#### Top20 DEG
```{r echo=F, warning=F, fig.width=10, fig.height=8}
Top20(sc.EdgeR$treat,30)
```

#### Volcano {.tabset}

##### treat
```{r echo=F, warning=F, message=F}
Volcano_plot(sc.EdgeR$treat, xscale = 4.5, yscale = 10, FC=log(FC), seuil=log(FC))
```

##### Glm
```{r echo=F, warning=F, message=F}
Volcano_plot(sc.EdgeR$glm, xscale = 4.5, yscale = 10, FC=log(FC), seuil=log(FC))
```

#### Heatmap {.tabset}

##### treat
```{r echo=F, fig.width=9, fig.height=11}
col <- colnames(sc.EdgeR$treat[['y']])
TopHeatmap(sc.EdgeR$treat, FDR=0.1, FC=0.5, cluster_cols = F)
```

##### Glm
```{r echo=F, fig.width=9, fig.height=11}
TopHeatmap(sc.EdgeR$glm, FDR=0.1, FC=0.5,cluster_cols = F)
```

#### Heatmap:Top {.tabset}

##### treat
```{r echo=F, fig.width=9, fig.height=11}
TopHeatmap(sc.EdgeR$treat, FDR=0.1, row = F, FC=0.5, cluster_cols = F)
```

##### Glm
```{r echo=F, fig.width=9, fig.height=11}
TopHeatmap(sc.EdgeR$glm, FDR=0.1, row = F, FC=0.5, cluster_cols = F)
```





## Single-Cell {.tabset}

### Liste

```{r}
item1 <- "Pregreffe"

singlet <- NormalizeData(singlet, assay = "RNA")
singlet <- ScaleData(singlet, features = rownames(singlet))
DE.Greffe.SC <- FindMarkers(singlet, ident.1 = item1, ident.2 = item2 , logfc.threshold = log2(FC), assay = "RNA", slot = "data")

DE.Greffe.SC.raw <- DE.Greffe.SC
save(DE.Greffe.SC.raw, file = "../Greffe/DE.Greffe.SC.raw.Rdata")

DE.Greffe.SC <- SC.Result(DE.Greffe.SC) ; DT.table(DE.Greffe.SC)
```

### Figure {.tabset}

#### PCA

```{r}
Seurat::DimPlot(object = singlet, group.by = "Condition", label.size = 5, pt.size = 1,reduction = "pca", label = T) & Seurat::NoLegend() & 
  xlab(label = paste0("PCA 1 : ", round(Seurat::Stdev(singlet[["pca"]])[1],2), " %")) & ylab(label = paste0("PCA 2 : ", round(Seurat::Stdev(singlet[["pca"]])[2],2), " %"))
```

#### Volcano

```{r, warning=FALSE}
SC.Volcano(DE.Greffe.SC, FC, yscale=350, seuil=0.25)
```

#### Heatmap

```{r, warning=FALSE, fig.width=9, fig.height=12}
DoHeatmap(singlet, features = DE.Greffe.SC$gene[1:200], group.by = "Condition", slot = "scale.data", assay = "integrated")
```



## Intersection {.tabset}

### Liste
```{r}
DE.intersect <- left_join(DE.Greffe.SC,DE.Greffe.PB2,by = "gene")
DE.intersect <- DE.intersect[DE.intersect$gene %in% intersect(DE.Greffe.SC$gene, DE.Greffe.PB2$gene),c("gene","p_val","PValue","logFC.x","logFC.y","FDR.x","FDR.y","description.y", "Expression.y")]
colnames(DE.intersect) <- c('Gene','SC.p.value','PB.p.value','SC.logFC','PB.logFC','SC.FDR','PB.FDR',"Expression.", "Description")

DT.table(DE.intersect)
```


### Venn Diagram {.tabset}

#### Treat
```{r, echo=FALSE, message=FALSE, error=FALSE}
x <- list(PseudoBulk = DE.Greffe.PB2$gene, SingleCell = DE.Greffe.SC$gene)
ggVennDiagram(x, category.names = c("PseudoBulk","SingleCell"), label_alpha = 1, label = "count") +
  scale_fill_gradient(low="white",high = "white") + theme(legend.position = "none") + ggtitle("DE list intersection") + theme(plot.title = element_text(hjust = 0.5))
```

#### Glm
```{r, echo=FALSE, message=FALSE, error=FALSE}
x <- list(PseudoBulk = DE.Greffe.PB.Glm2$gene, SingleCell = DE.Greffe.SC$gene)
ggVennDiagram(x, category.names = c("PseudoBulk","SingleCell"), label_alpha = 1, label = "count") +
  scale_fill_gradient(low="white",high = "white") + theme(legend.position = "none") + ggtitle("DE list intersection") + theme(plot.title = element_text(hjust = 0.5))
```




# Enrichissement {.tabset}

## Pseudo-Bulk - les databases (union DB papiers) {.tabset}
```{r}
HM.Path<- gmtPathways("../other/gmt/h.all.v7.5.1.symbols.gmt")
test   <- sc.EdgeR$treat
table  <- test$tr$table
ranks  <- table$logFC ; names(ranks) <- rownames(table)
idx    <- ids2indices(HM.Path,identifiers=table$genes)
```

### Liste
```{r}
HM.FGSEA.Result <- fgsea(pathways = HM.Path, stats = ranks, minSize = 5) ; Path <- c(HM.FGSEA.Result[ES > 0][head(order(pval), n=10), pathway], rev(HM.FGSEA.Result[ES < 0][head(order(pval), n=10), pathway]))

HM.FGSEA.Result.raw <- HM.FGSEA.Result
save(HM.FGSEA.Result.raw, file = "../Greffe/HM.FGSEA.Result.raw.Rdata")

HM.FGSEA.Result <- HM.FGSEA.Result %>% as_tibble() %>% dplyr::arrange(desc(abs(NES))) %>% dplyr::select(-ES) %>% dplyr::arrange(padj) %>% dplyr::mutate_if(is.double, round, 7) %>% rename(Hallmark = pathway) 
HM.FGSEA.Result <- HM.FGSEA.Result[HM.FGSEA.Result$padj<0.05,]
DT.table(HM.FGSEA.Result)
```

### EnrichPlot {.tabset}
```{r, results='asis'}
for(i in 1:length(HM.FGSEA.Result$Hallmark)){
  cat("#### ", stringr::str_replace(HM.FGSEA.Result$Hallmark[i],"HALLMARK_",""), "\n\n")
  print(plotEnrichment(HM.Path[[HM.FGSEA.Result$Hallmark[i]]], ranks) + labs(title=HM.FGSEA.Result$Hallmark[i])) ; cat("\n\n")
}
```


## Single-cell
```{r}
ES <- enrichIt(obj = singlet, gene.sets = GS, groups = 1000, cores = 10, min.size = 5)
save(ES, file = "../Greffe/ES.RDs")
load(file = "../Greffe/ES.RDs")
singlet <- AddMetaData(singlet, ES)
singlet@meta.data$active.idents <- singlet@active.ident

HM <- data.frame(singlet[[colnames(ES)]], Idents(singlet)) ; colnames(HM)[ncol(HM)] <- "Condition"
HM.Result <- getSignificance(HM, group = "Condition", fit = "Wilcoxon") 
HM.Result <- HM.Result[order(HM.Result$FDR),]
HM.Result <- HM.Result[HM.Result$FDR<0.05,]

HM.SC.Result0.raw <- HM.Result
save(HM.SC.Result0.raw, file = "../Greffe/HM.SC.Result0.raw.Rdata")


HM.SC.Result <- SC.Enrichissement(singlet, ES, condition) ; 
DT.table(HM.SC.Result)
```





## Intersection {.tabset}

### Liste
```{r}
HM.intersect <- left_join(HM.FGSEA.Result,HM.SC.Result,by = "Hallmark") 
HM.intersect <- HM.intersect[HM.intersect$Hallmark %in% intersect(HM.FGSEA.Result$Hallmark, HM.SC.Result$Hallmark),c("Hallmark","p.value","FDR","pval","padj", "leadingEdge")] ; colnames(HM.intersect) <- c('Hallmark','SC.p.value','SC.FDR', 'PB.p.value','PB.FDR', 'PB.genes')

DT.table(HM.intersect)
```

### Diagram
```{r, echo=FALSE, message=FALSE, error=FALSE}
x <- list(PseudoBulk = as.vector(HM.FGSEA.Result$Hallmark), SingleCell = as.vector(HM.SC.Result$Hallmark))
ggVennDiagram(x, category.names = c("PseudoBulk","SingleCell"), label_alpha = 1, label = "count") +
  scale_fill_gradient(low="white",high = "white") + theme(legend.position = "none") + ggtitle("Hallmark list intersection") + theme(plot.title = element_text(hjust = 0.5))
```
